Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Comments on 7 and 8

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then controlled by the activation that is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals that it does not enforce confidentiality (either stepwise or
observational): it violates of the temporal aspect of confidentiality in
corner cases where data can leak across calls to the same callee.  We
propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)}

@@

\leo{I am not sure we've hit the correct level of detail in this section.}

In this section we detail an existing enforcement mechanism,
Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
the formal stack-safety property described in the previous section; in
fact, it enforces a stronger, inductive version of that property which strengthens the
local state encapsulation component by checking for integrity and
confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}

@@

\leo{I'm not a fan of the next paragraph.}
Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
For example, the program in \cref{fig:passingsimple} manages the stack
in this way. Here, the call pseudo-instruction to {\tt f} in line 5
hides a blessed sequence of instructions that allocate space to save
its return address to {\tt main} and seal the protected part of the
stack before {\tt f}'s body starts executing at 10. Conversely, the
return instruction in line 15 restores and jumps to the return address
and deallocates its frame space, and unseals the corresponding parts
of the stack. The function itself allocates and deallocates its own
private frame space in instructions 10 and 14. Additionally, unlike in
the simplified presentation of the policy, the caller shares the
arguments passed to the callee through the stack, and for this creates
and initializes (3--4) and destructs (6) the argument part of the
callee's stack frame.
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

@@

We validate the micro-policy of the previous section by systematically
testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
extended it with a runtime monitor implementing the stack-safety
micro-policy. \leo{Footnote: code will be made public} We chose the
Coq proof assistant as the setting for our implementation for two
reasons: first, to ensure that our coinductive trace definitions are
well-formed, and reason about them\leo{Do we prove anything
  interesting anymore?}, and to leverage the power of the QuickChick
property-based testing framework.\leo{cite}

To use QuickChick, we built generators that produce:
\begin{enumerate}
\item An initial RISC-V machine state, including most notably
  the program to be executed.
\item An initial policy state, tagging instructions corresponding to
  ``blessed'' call or return sequences appropriately, while marking
  all potential stack locations as $\tagNoDepth$.
\item A code map and a stack map, accurately characterizing the behavior
  of callers and callees.
\end{enumerate}
%
That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also policy
meta-information and code annotations that would normally be produced
by the compiler.

\paragraph*{Generator Overview}

To write such generators we build on the work of \leo{ICFP 2013},
where they introduced {\em generation by execution} to produce
interesting programs. Generation by execution receives as an input a
partially instantiated machine state, and attempts to generate an
instruction (or a sequence of instructions such as a ``blessed''
sequence) that makes sense locally (e.g., jumps to a potentially valid
code location or loads from a potentially valid stack location). Then,
we step the machine and repeat the process until we generate or
execute some target number of instructions (or if we reached a point
where the machine can't step anymore).

We extend this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we sometimes allow the generation to relax those constraints,
introducing potentially ill-formed flows: that causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.

@@

To evaluate the effectiveness of our testing, we turn to {\em mutation
  testing}\leo{cite}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest
(here, stack safety) to fail, and ensure that the testing framework
can find them. The bugs we used for our evaluation were both
artificially generated (by deliberately weakening the micro-policy in
ways that should break enforcement), or actual bugs that we discovered
through testing in our implementation. We elaborate on some such bugs
below.

The mean-time-to-failure (MTTF) for various bugs can be found in
\leo{Table with numbers?}. All experiments were run in a \leo{Machine
  spec}. \leo{Something like: There is really no baseline to compare
  this against, but include it here for the future/reasons?}

The artificially injected bugs focus on weakening the tag propagation
and checking mechanism. \leo{Mention them all?}
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the one of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
that restriction by not checking the depth equality, or not checking
the tags at all, causing bugs of different difficulty.
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing these constraints also causes violations to the
property.

Using our testing framework, we also found various errors in our
implementation, ranging from trivial typos to ones that required an
intriguingly complex setup to appear. Most notably, the initial
``blessed'' call sequence/policy combination did not initialize all
locations for the newly allocated stack frame correctly, but left some
of them as $\tagNoDepth$. This allows for a potential integrity
violation, but only if a rather complicated sequence of events
occured: \leo{call - write - call (seal) - return - call - write?}

>>>>

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then controlled by the activation that is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
stepwise or
observational): it violates of the temporal aspect of confidentiality in
corner cases where data can leak across calls to the same callee.  We
propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)}

@@

\bcp{``Enforcement'' seems like too broad a title.  What about calling it
  ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
  Safety?''? And what about folding the testing section into this one?  That
would keep related things together, would avoid appearing to make overly
grand claims in either section, and would give us an excuse to write less
about the mechanism here---we only need to write enough about the mechanism
itself that readers can understand our validation case study.}

\leo{I am not sure we've hit the correct level of detail in this
  section.}\bcp{Agreed!}

In this section we sketch an existing enforcement mechanism,
Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
the formal stack-safety property described in the previous section; in
fact, it enforces a stronger, inductive version of that property which strengthens the
local state encapsulation component by checking for integrity and
confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}

@@

\leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
  the weeds.  I don't think this level of detail is needed, in the context
  of a paper that is mostly about our definitions, not theirs.}
Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
For example, the program in \cref{fig:passingsimple} manages the stack
in this way. Here, the call pseudo-instruction to {\tt f} in line 5
hides a blessed sequence of instructions that allocate space to save
its return address to {\tt main} and seal the protected part of the
stack before {\tt f}'s body starts executing at 10. Conversely, the
return instruction in line 15 restores and jumps to the return address
and deallocates its frame space, and unseals the corresponding parts
of the stack. The function itself allocates and deallocates its own
private frame space in instructions 10 and 14. Additionally, unlike in
the simplified presentation of the policy, the caller shares the
arguments passed to the callee through the stack, and for this creates
and initializes (3--4) and destructs (6) the argument part of the
callee's stack frame.
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

@@

We validated the micro-policy of the previous section by systematically
testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
extended it with a runtime monitor implementing the stack-safety
micro-policy.\leo{Footnote: code will be made public}\bcp{For the reviewers?
 Or later?} We chose the
Coq proof assistant as the setting for our implementation for two
reasons: first, to ensure that our coinductive trace definitions are
well-formed, and reason about them\leo{Do we prove anything
  interesting anymore?}, and to leverage the power of the QuickChick
property-based testing framework~\citep{Pierce:SF4}.

To use QuickChick, we built random test-case generators that produce:
\begin{enumerate}
\item An initial RISC-V machine state, including most notably
  the program to be executed.
\item An initial policy state, tagging instructions corresponding to
  ``blessed'' call or return sequences appropriately, while marking
  all potential stack locations as $\tagNoDepth$.
\item A code map and a stack map, accurately characterizing the behavior
  of callers and callees.
\end{enumerate}
%
That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also the policy
meta-information and code annotations that would normally be produced
by the compiler.

\paragraph*{Generators}

To write such generators we build on the work of \leo{ICFP 2013},
where they introduced {\em generation by execution} to produce
interesting programs. Generation by execution receives as an input a
partially instantiated machine state and attempts to generate an
instruction (or a sequence of instructions such as a ``blessed''
sequence) that makes sense locally (e.g., jumps go to a potentially valid
code location and loads read from from a potentially valid stack
location). Then
we step the machine and repeat the process until we generate or
execute some target number of instructions, or if we reach a point
where the machine can't step anymore.

We extended this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we sometimes allow the generation to relax those constraints,
introducing potentially ill-formed flows: that causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.

@@

To evaluate the effectiveness of testing against our formal properties, we
used {\em mutation testing}\leo{cite}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest (here,
stack safety) to fail, and ensure that the testing framework can find
them. The bugs we used for our evaluation were either artificially generated
by us (deliberately weakening the micro-policy in ways that we expected
should break its guarantees), or actual bugs that we discovered through
testing our implementation. We elaborate on some such bugs below.

The mean-time-to-failure (MTTF) for various bugs can be found in \leo{Table
  with numbers?}\bcp{Sure.}. All experiments were run in a \leo{Machine
  spec}. \leo{Something like: There is really no baseline to compare this
  against, but include it here for the future/reasons?}

The artificially injected bugs focus on weakening the tag propagation
and checking mechanism. \leo{Mention them all?}\bcp{Nah, especially if they
  are all listed in the table.}
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the one of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
that restriction by not checking the depth equality, or not checking
the tags at all, causing bugs of different difficulty.
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing these constraints also causes violations to the
property.

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code\bcp{right?}; we re-implemented it in Coq targeting
RISC-V).  These errors ranged from trivial typos to
ones that required an 
intriguingly complex setup to appear.  Notably, our first try at the 
``blessed'' call sequence/policy combination did not initialize all
locations for the newly allocated stack frame correctly, but left some
of them as $\tagNoDepth$. This allowed for a potential integrity
violation, but only if a rather complicated sequence of events
occured: \leo{call - write - call (seal) - return - call - write?}
